---
title: "Room Costs in Norwegian Counties"
author: "Lucas Stark"
date: "18 9 2019"
output: 
  html_document:
    number_sections: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(scales)
library(reshape2)
library(Hmisc)
rversion <- paste0(R.Version()[["major"]],".",R.Version()[["minor"]])
```

## Introduction

```{r load, include=FALSE}
#Load Data
url1 <- "http://data.ssb.no/api/v0/dataset/95274.csv?lang=no"
url2 <- "http://data.ssb.no/api/v0/dataset/95276.csv?lang=no"
dat1 <- read.csv(url1, sep=";", dec = ",")
dat2 <- read.csv(url2, sep=";", dec = ",")
```

The data for this project was taken from the webpage of the [statistisk sentralbyrå](http://data.ssb.no/api/v0/dataset/) of Norway. Two datasets about prices and capacities of rooms for overnight stays were downloaded into R (Version `r rversion`). After this, the dataframes were merged and cleaned. Eventually, a difference score between the individual mean values of the Norwegian counties' price per room and the mean of the overall mean of Norway was computed.

```{r rename, include=FALSE}
#Rename Columns
colnames <- c("region", "date", "variable", "value")
colnames(dat1) <- colnames
colnames(dat2) <- colnames
```

```{r convert, include=FALSE}
#Turn Characters to Date Variables
dat1 <- dat1 %>%
  mutate(date = gsub("M","01",date)) %>%
  mutate(date = lubridate::ydm(date))
dat2 <- dat2 %>%
  mutate(date = gsub("M","01",date)) %>%
  mutate(date = lubridate::ydm(date))
```

```{r recode, include=FALSE}
# Recode variable name
recode_task1 <- function(x){
   y <- recode(x, 
                           "Kapasitetsutnytting av rom (prosent)" = 
                           "Rooms percentage capacity utilization",
                           "Kapasitetsutnytting av senger (prosent)" =
                           "Beds percentage capacity utilization",
                           "Pris per rom (kr)" =
                           "Price per room")
   y
}
dat1 <- dat1 %>%
  mutate(variable = recode_task1(variable))
dat2 <- dat2 %>%
  mutate(variable = recode_task1(variable))

dat1$value[dat1$value==0] <- NA
dat2$value[dat2$value==0] <- NA
```

```{r merge, include=FALSE}
dat1 <- dat1 %>%
  mutate(region = as.character(region),
         value = as.numeric(value))
dat2 <- dat2 %>%
  mutate(region = as.character(region),
         value = as.numeric(value))
dat <- merge(dat1, dat2, by=c("date","variable"))
```

## Analysis

### Least and most expensive counties and years

```{r newvariable, include=FALSE}
dat <- dat %>%
  mutate(value.diff = value.x - value.y)
means_county <- dat %>%
  filter(variable=="Price per room") %>%
  group_by(region.x) %>%
  summarise(mean = mean(value.diff, na.rm=T))
means_years <- dat %>%
  filter(variable=="Price per room") %>%
  group_by(lubridate::year(date)) %>%
  summarise(mean = mean(value.diff, na.rm=T))
colnames(means_years) <- c("year","mean")
overall_mean <- mean(dat2[dat2$variable=="Price per room","value"],na.rm=T)
max_county <- gsub("\\d", "",means_county[means_county$mean==max(means_county$mean),"region.x"])
min_county <- gsub("\\d", "",means_county[means_county$mean==min(means_county$mean),"region.x"])
#die hier gehen nich so??
max_year_vec <- means_years$mean==max(means_years$mean, na.rm=T)
max_year_vec[is.na(max_year_vec)] <- F
max_year <- means_years[max_year_vec,"year"]
min_year_vec <- means_years$mean==min(means_years$mean, na.rm=T)
min_year_vec[is.na(min_year_vec)] <- F
min_year <- means_years[min_year_vec,"year"]
```

Two aspects of the dataframe were analyzed to get an overview over the relative price situation inbetween Norwegian counties. 

First, a mean value was computed for each county over every year of data observation. That made apparent that `r max_county` was the most expensive county with `r round(max(means_county$mean),2)` NOK per room above the national average. `r min_county` was the least expensive county with `r round(min(means_county$mean),2)` NOK below the national average.

Second, a mean value was computed for each year over every county. That made apparent that `r max_year` was the most expensive year with `r round(means_years$mean[max_year_vec],2)` NOK per room above the average of all years. `r min_year` was the least expensive year with `r round(means_years$mean[min_year_vec],2)` NOK below the average.

### Troms' Prices per Room over Time from 1999 till Today

As a next step, the monthly averages of prices per room for the Troms county were checked in comparison to the national average.

```{r, echo=FALSE, warning = FALSE}
troms_graphic <- dat %>%
  filter(variable=="Price per room") %>%
  filter(region.x=="19 Troms - Romsa") %>%
  ggplot(aes(x=date, y=value.diff)) +
  geom_line(col="blue") +
  geom_smooth(method = 'loess',formula = 'y ~ x')  +
  ggtitle("Troms' Price per Room in Relation to National Average") + 
  xlab("Date") + 
  ylab("Price per Room compared to National Average") + 
  scale_x_date(labels = date_format("%Y"), breaks='2 years', 
               limits = c(as.Date("1999-01-01"),Sys.Date()))
troms_graphic

```

The graph shows that the room prices in the county of Troms were mostly below average after 1999. However, since 2016 the prices started to variate more extremely. 

### Relationship between Room Capacity and Price since January 2010

Per county, is there any relationship between room capacity and price since January 2010? For this, the correlations for each county were computed. These can be seen in the following table.

```{r, echo = F}

data_wide <- dcast(dat, date + region.x ~ variable, value.var="value.x")
names(data_wide)[3:5] <- c("roomCap","bedCap","priceRoom")

cor_county <- data_wide %>% 
  filter(date >= as.Date("2010-01-01")) %>%
  group_by(region.x) %>%
  mutate(cor = rcorr(x=roomCap, y=priceRoom,
            type="pearson")[["r"]][2],
         p = rcorr(x=roomCap, y=priceRoom,
            type="pearson")[["P"]][2])
cortable <- cor_county %>%
  filter(date == as.Date("2010-01-01")) #just choose one
cortable <- cortable[,c(2,6:7)]  
for (i in 1:length(cortable$cor)){
  cortable[i,1] <- gsub(c("\\d"), "", cortable[i,1])
  cortable[i,1] <- gsub(c("\\(-\\)"), "", cortable[i,1])
}
cor_sign <- ifelse(cortable[,"p"]<.05,T,F)
mean_cor_sign <- mean(cortable$cor[cor_sign], na.rm=T)
names(cortable) <- c("County","Correlation","P value")
knitr::kable(cortable, format="pandoc", digits=2)

```

It is observable that `r ifelse(mean(cor_sign,na.rm=T)>.05, print("more"), print("less"))` than half of the correlations are significant with a significance level of $\alpha < .05$. The average of these significant correlations is `r round(mean_cor_sign,2)`. This speaks for an overall `r ifelse(mean_cor_sign>0,print("positive"),print("negative"))` relationship between room capacity and price in these counties.
